---
title: "Review of Lakpa Sherpa’s R Assignment"
author: "Modestos A. Modestou"
date: "2025-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hello Lakpa,

Overall, excellent job! I have a few suggestons for you:

1. When I ran your R Markdown file the first time, I couldn't run it because of the command "View (snp_position)" on  line # 53. The View() command cannot be used inside an R Markdown file when knitting. So I will sugggest either removing it from the R Markdown file (you don't really need it since you include the command "head (snp_position)") or use the command "if (interactive()) View(snp_position). This command ensures that View() only runs in the RStudio console and not during knitting.

2. You can modify your first chunk for data visualization, as shown below, to have the chromosomes appropriately numbered on the x-axis. By the way you can name your chunks for easy reference as {r chunk_name}.

```{r}

# Format Chromosomes: Pad numbers with leading zeros, but keep "unknown" and "multiple"

maize_snp$Chromosome <- ifelse(maize_snp$Chromosome %in% c("unknown", "multiple"),
                           maize_snp$Chromosome,
                           sprintf("%02d", as.numeric(maize_snp$Chromosome)))
teosinte_snp$Chromosome <- ifelse(teosinte_snp$Chromosome %in% c("unknown", "multiple"),
                              teosinte_snp$Chromosome,
                              sprintf("%02d", as.numeric(teosinte_snp$Chromosome)))

# Calculate the number of SNPs per chromosome for maize and teosinte
maize_snps_per_chr <- maize_snp %>%
  group_by(Chromosome) %>%
  summarize(n = n())

teosinte_snps_per_chr <- teosinte_snp %>%
  group_by(Chromosome) %>%
  summarize(n = n())

# Combine the data frames for maize and teosinte, adding a Group column
snps_per_chr <- bind_rows(
  mutate(maize_snps_per_chr, Group = "Maize"),
  mutate(teosinte_snps_per_chr, Group = "Teosinte")
)

# Dot plot to visualize SNP counts per chromosome
ggplot(snps_per_chr, aes(x = factor(Chromosome), y = n, color = Group)) +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  labs(title = "Number of SNPs per Chromosome (Dot Plot)",
       x = "Chromosome", y = "Number of SNPs") +
  theme_minimal() +
  scale_color_manual(values = c("Maize" = "blue", "Teosinte" = "green"))

#From graph, chromosome 1 has the highest SNP where as the chromosome 10 has lowest SNPs.
```

3. I have the same graphs for the SNP numbers per chromosome and the distribution of SNPs across chromosomes. I think, however, it doesn't make sense that the number and distribution of SNPs for maize and teosinte are precisely the same. They shouldn't; maize likely has many more SNPs than its ancestor, teosinte. Unfortunately, I can't offer any suggestions here. I am still struggling myself to figure out how to fix it. 

4. The last thing I want to mention is that you may want to include the intermediate files you generated to get to the two final output files for maize and teosinte.
